---
title: "ISYE6501 HW5"
author: "Keh-Harng Feng"
date: "June 15, 2017"
header-includes:
    - \usepackage{placeins}
output: 
  bookdown::html_document2:
    fig_caption: TRUE
    toc: FALSE
urlcolor: blue
---
```{r setup, include=FALSE}
library('knitr')
library('caret')
library('glmnet')
library('glmnetUtils')
library('parallel')

opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, tidy = TRUE, cache = TRUE)
options(digits = 4)
```

## Preface
This is a reproducible report with most of the codes doing the heavy lifting hidden in the background. You can download the source code of the report by [clicking here](https://github.com/fengkehh/ISYE6501WK4/blob/master/HW4_Report.Rmd). All of the information you need as a reivewer is in the report, not in the code. You should NEVER run any R scripts from an untrusted source on your own computer.  

# Question 1
**Using the crime data set from Homework 3, build a regression model using:**
1. **Stepwise regression**
2. **Lasso**
3. **Elastic net**
**For Parts 2 and 3, remember to scale the data first – otherwise, the regression coefficients will be on different scales and the constraint won’t have the desired effect.**
**For Parts 2 and 3, use the glmnet function in R.**

```{r}
q1_data <- read.table('uscrime.txt', header = TRUE)

q1_data$So <- factor(q1_data$So)

n <- nrow(q1_data)

set.seed(123)
inTrain <- sample(1:n, size = ceiling(n*0.9))

q1.train <- q1_data[inTrain,]
q1.test <- q1_data[-inTrain,]

# Hack
q1.test <- rbind(q1.train[1,], q1.test)
q1.test <- q1.test[-1,]

# Preprocessing
#preProc1 <- preProcess(q1.train[,-c(2, 16)], method = 'corr')
#q1.train.preproc1 <- predict(preProc1, q1.train)
preProc <- preProcess(q1.train.preproc1[,-16], method = c('BoxCox', 'center', 'scale'))
q1.train.preproc2 <- predict(preProc2, q1.train.preproc1)
```

Variable `So` is converted to a 2-level factor as it is categorical. The data is then splitted into a training (~90%) and testing (~10%) sets as usual. Predictor variables from the training set are transformed using Box-Cox transformation then centered and scaled (see my [HW3 Report](https://github.com/fengkehh/ISYE6501WK3/blob/master/HW3_Report.pdf)). The response `Crime` is not preprocessed.

## Part 1: Stepwise Regression
Stepwise regression in both directions is carried out using the `step()` function and AIC as the selection metric. The initial starting point is an OLS model using all features. 

```{r}
# Initial Model for Stepwise Selection
model.init <- lm(Crime ~ ., data = q1.train.preproc2)

# Stepwise Selection
model.stepboth <- step(model.init, direction = 'both')
```
